---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Jessica Im
thanks: "Code and data are available at: https://github.com/jimessica/boycott"
date: today
date-format: long
abstract: "Introduction to the topic and its importance. What I did in my research and analysis. What was found from my research and analysis. Significance of my findings."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(tinytex)
library(kableExtra)
library(knitr)
library(janitor)
library(kableExtra)
library(ggplot2)
library(xts)


cleaned_data_NKE <- read.csv("../data/analysis_data/cleaned_data_NKE.csv")
cleaned_data_NKE <- 
  cleaned_data_NKE |>
  mutate(Date = lubridate::ymd(Date))
cleaned_data_NKE$Date <- 
  as.Date(cleaned_data_NKE$Date, format='%Y-%m-%d')

cleaned_data_PRKS <- read.csv("../data/analysis_data/cleaned_data_PRKS.csv")
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

In [Background], the history of why each company was or is being boycotted is explained. In the following section, [Data], 

# Background

## Nike (NKE)

In August of 1992, Jeffrey Ballinger published “The new free-trade heel: Nike's profits jump on the backs of Asian workers” in a Harper’s Bazaar issue [@harpers]. This publication revealed the low pay and poor working conditions in Nike’s Southeast Asian and Middle Eastern sweatshops [@ethically]. Later that year, Nike released the “Code of Conduct and Memorandum of Understanding” for contract workers, but the labour controversy was well underway. In the following years, protests in Southeast Asia and the United States raged on. Protests at the Barcelona Olympics in 1992 and the founding of Ballinger’s labour group Press for Change in 1993 brought new waves of media attention to the cause. After years of Nike promotions fueling more public outrage, in 1998, Nike’s then-CEO Phil Knight announced that Nike will raise minimum wage, increase monitoring, and introduce U.S. clean air standards at all factories [@ethically]. Knight acknowledged the brand’s association with workplace abuse and stated that "the American consumer doesn’t want to buy products made under abusive conditions”. This speech marked the shift away from boycotting and for the purposes of this paper will represent the end of Nike boycotting. 


## United Parks and Recreation
Blackfish is a film focused on the cruel captivity of orcas by focusing on the infamous killer whale, Tilikum, who was kept in captivity at SeaWorld and was involved in the death of 3 people [@WAP]. The film was released in 2013 and shown on major channels like CNN, Netflix, and Universal Studios around the world and over its 14-week box office period in North America, earned $2,073,582 [@wikiBlackfish]. In March of 2014, California Assemblyman Richard Bloom proposed that orcas are banned from being held captive for entertainment [@CNNorcaban]. Later that year, U.S. Congressmen Adam Schiff and Jared Huffman amended the Agriculture Appropriations Act to require the USDA’s Animal & Plant Health Inspection Service to update and finalize regulations on marine mammal captivity [@SchiffandHuffman]. On behalf of SeaWorld investors, the Rosen Law Firm filed a class action lawsuit against SeaWorld in 2014 for 



## Chevron

## Starbucks
On October 9th the Starbucks Workers United posted to their Twitter “Solidarity with Palestine” [@apnews]. This tweet was up for 40 minutes before it was taken down and was originally posted without the approval of union leaders according to Workers United. Even after the post was taken down, posts and retweets from local Starbucks United twitter accounts in support of Palestine and condemning Israel were circulating. On October 13th, Starbucks sent a letter to Workers United, demanding the union to stop using the Starbucks name and similar logo. Workers United retaliated saying that Starbucks was using the tragedy in the Middle East to support their anti-union campaign. 


# Data {#sec-data}

In @tbl-rawdataNKE, the data's date and stock price are shown.

```{r}
#| label: tbl-rawdataNKE
#| tbl-cap: "Description of Raw Nike Stock Variables"
#| echo: false

rawdataNKE <- data.frame( 
  "Variables" = c( "Date", "Close"),
  "New Name" = c( "Date", "Stock Price"), 
  "Description of Variables" = c("The day of the stock price", "The value of the stock at market close"), 
  "Example Value" = c("1991-07-12", "1.355469")
)

knitr::kable(rawdataNKE, 
      col.names = c("Variables", 
                    "New Name",
                    "Description of Variables" ,
                    "Example Value"),
      booktabs = TRUE,
      linesep = " "
      )
```

@fig-linegraphNKE shows the change in stock price during boycotting of Nike.
```{r}
#| label: fig-linegraphNKE
#| fig-cap: Changes in Nike Stock Price from One Week Before the Commencement of Labour Boycotts to One Week After Their Pledge to End Poor Working Conditions and Child Labour
#| echo: false

cleaned_data_NKE |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "July 11th, 1991 to May 19th, 1998", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "12 months",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(0,10) + 
  theme_minimal()

```

@fig-linegraphPRKS shows the change in stock price during boycotting of United Parks and Recreation.
```{r}
#| label: fig-linegraphPRKS
#| fig-cap: Changes in United Parks and Recreation Stock Price from One Week Before the Commencement of Blackfish was Released to One Week After Their Pledge to End Orca Breeding
#| echo: false

cleaned_data_PRKS |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "July 12th, 2013 to May 17th, 2016", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "12 months",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(15,40) + 
  theme_minimal()

```



```{r}
#| label: fig-linegraphNIKE
#| fig-cap: Changes in United Parks and Recreation Stock Price from One Week Before the Commencement of Blackfish was Released to One Week After Their Pledge to End Orca Breeding
#| echo: false

```

Talk way more about it.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]


```

## Diagnostics

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2


```

\newpage

# References
