---
title: "Boycotts require significant deterioration in a company's public image to have an impact on their stock prices."
subtitle: "An analysis of the stocks of 3 large U.S. companies boycotted between 1992 to 2024."
author: 
  - Jessica Im
thanks: "Code and data are available at: [https://github.com/jimessica/boycott)](https://github.com/jimessica/boycott)"
date: today
date-format: long
abstract: "Introduction to the topic and its importance. What I did in my research and analysis. What was found from my research and analysis. Significance of my findings."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(tinytex)
library(kableExtra)
library(knitr)
library(janitor)
library(kableExtra)
library(ggplot2)
library(xts)
library(tidyquant)
library(arrow)
library(readr)
library(gridExtra)

## Import cleaned data and turn "Date" into class Date
# Nike
cleaned_data_NKE <- read.csv("../data/analysis_data/cleaned_data_NKE.csv")
cleaned_data_NKE <- 
  cleaned_data_NKE |>
  mutate(Date = lubridate::ymd(Date))
cleaned_data_NKE$Date <- 
  as.Date(cleaned_data_NKE$Date, format='%Y-%m-%d')

# United Parks and Recreation
cleaned_data_PRKS <- read.csv("../data/analysis_data/cleaned_data_PRKS.csv")
cleaned_data_PRKS <- 
  cleaned_data_PRKS |>
  mutate(Date = lubridate::ymd(Date))
cleaned_data_PRKS$Date <- 
  as.Date(cleaned_data_PRKS$Date, format='%Y-%m-%d')

#Starbucks
cleaned_data_SBUX <- read.csv("../data/analysis_data/cleaned_data_SBUX.csv")
cleaned_data_SBUX <- 
  cleaned_data_SBUX |>
  mutate(Date = lubridate::ymd(Date))
cleaned_data_SBUX$Date <- 
  as.Date(cleaned_data_SBUX$Date, format='%Y-%m-%d')


## Import S&P 500 Index and Dow Jones Index Data
# S&P 500
cleaned_data_SPX <-read_parquet("../data/analysis_data/cleaned_data_SPX.parquet")

# Dow Jones 
cleaned_data_DJI <-read_parquet("../data/analysis_data/cleaned_data_DJI.parquet")

# Nasdaq 100
cleaned_data_NDX <-read_parquet("../data/analysis_data/cleaned_data_NDX.parquet")


```

# Introduction
Boycotting is a common way for consumers to cast their vote on the amount of power corporations and institutions should have. Boycotting is usually done by avoiding purchasing from a company or withdrawing support for an institution [@ethicalconsumer]. It is oftentimes a response to an ethical violation and has had most success by impacting the public image of the company through generating negative media coverage [@northwestern]. As businesses have grown and markets have been dominated by oligopolies, consumers feel ignored and alienated [@kokuthreats]. As a result, boycotting has increased and between 1984 to 1990 boycotting increased by 769%. Recently, the Boycott, Divestment, Sanctions (BDS) movement in support of Palestine’s freedom and justice has introduced a new wave of companies being boycotted. Many debates surround the effectiveness of boycotting and whether it can inflict substantial financial losses for  targeted organizations. Levesque and Nam found that the impacts of boycotts on stock price is highly varied and the overall stock market was not significantly impacted by boycotts (2019). In contrast, Pruitt and Friedman found that target firms suffered from significantly decreased stock prices following boycotts at 21 companies of interest (1986). Many factors may be involved, but researchers have had a hard time pinpointing which ones are the most impactful; Pruitt and Friedman failed to identify any significant relationships between the target’s damage and boycott attributes (1986). According to Brayden King, a boycott is most effective when it generates a lot of negative media coverage [@northwestern]. Levesque and Nam found that companies with a bad reputation before the boycott, larger market capital, and frequent past scandals were more likely to be impacted by boycotts (2019). Overall, the particular factors related to the impact of boycotts and whether boycotts have a significant financial punishment are unclear in existing literature.

The objective of this paper is to add to existing literature by determining the impact of boycotting on stock prices and the factors involved in its effectiveness, particularly media coverage. The estimand is the change in stock price during the boycott period. The beginning of the boycott is determined by an event which sparks the most media coverage or is mentioned the most in the media. The end of the boycott is determined by mention in the media of a “turning point” for the company where the reason for the boycott is addressed. Since Starbucks is an ongoing boycott at the time of this research, the “end date” is the date research began. The stock prices when the market closed were arbitrarily selected instead of the stock price at open. To investigate, Nike, United Parks and Recreation, and Starbucks are analyzed as large U.S. companies that were famously boycotted during the desired time period. The analysis includes a detailed view of the difference in stock prices overall, during periods of heavy media coverage, and in comparison with market trends. From analysis, the stock prices during the boycott of Nike were found to not be impacted, United Parks and Recreation was heavily impacted, and Starbucks were mildly impacted. The range in impact can be attributed to overall company image, product type, earnings calls, unionization, and social media. This result supports the importance of media coverage, but also the lack of financial repercussions to many boycotted companies. 


In [Background](#sec-background), the history of why each company was or is being boycotted is explained. In the following section, [Data](#sec-data), the dataset, variables used, limitations of the data, and methodology for analysis and cleaning are explained. In [Results](#sec-results), there are detailed explorations of the difference in stock price over the boycott, comparisons with the overall market, and an analysis of important dates during the boycotts. Finally, the [Discussion](#sec-discussion) explores the impact of earnings calls, product or service type, and social media. 

# Background {#sec-background}

## Nike (NKE)

In August of 1992, Jeffrey Ballinger published “The new free-trade heel: Nike's profits jump on the backs of Asian workers” in a Harper’s Bazaar issue [@harpers]. This publication revealed the low pay and poor working conditions in Nike’s Southeast Asian and Middle Eastern sweatshops [@ethically]. Later that year, Nike released the “Code of Conduct and Memorandum of Understanding” for contract workers, but the labour controversy was well underway. In the following years, protests in Southeast Asia and the United States raged on. Protests at the Barcelona Olympics in 1992 and the founding of Ballinger’s labour group Press for Change in 1993 brought new waves of media attention to the cause. In April 1997 a boycott was called for by the Council on American-Islamic Relations due to a logo which resembled the word "Allah" in Arabic script [@islamiccouncil]. Two months later Nike announced that they would cease to sell these shoes and apologized for the unintentional offense [@nikepulled]. In October of 1997, a non-profit organization called Global Exchange used the internet to organize the first global protest against Nike [@carty]. 84 communities in 12 different countries participated in this anti-sweatshop activism [@carty]. After years of Nike promotions fueling more public outrage, in 1998, Nike’s then-CEO Phil Knight announced that Nike will raise minimum wage, increase monitoring, and introduce U.S. clean air standards at all factories [@ethically]. Knight acknowledged the brand’s association with workplace abuse and stated that "the American consumer doesn’t want to buy products made under abusive conditions”. This speech marked the shift away from boycotting and for the purposes of this paper will represent the end of Nike boycotting.

## United Parks and Recreation

Blackfish is a film focused on the cruel captivity of orcas by focusing on the infamous killer whale, Tilikum, who was kept in captivity at SeaWorld and was involved in the death of 3 people [@WAP]. The film was released in 2013 and shown on major channels like CNN, Netflix, and Universal Studios around the world and over its 14-week box office period in North America, earned \$2,073,582 [@wikiBlackfish]. In March of 2014, California Assemblyman Richard Bloom proposed that orcas are banned from being held captive for entertainment [@CNNorcaban]. Later that year, U.S. Congressmen Adam Schiff and Jared Huffman amended the Agriculture Appropriations Act to require the USDA’s Animal & Plant Health Inspection Service to update and finalize regulations on marine mammal captivity [@SchiffandHuffman]. On behalf of SeaWorld investors, the Rosen Law Firm filed a class action lawsuit against SeaWorld in 2014 for

## Starbucks

On October 9th the Starbucks Workers United posted to their Twitter “Solidarity with Palestine” [@apnews]. This tweet was up for 40 minutes before it was taken down and was originally posted without the approval of union leaders according to Workers United. Even after the post was taken down, posts and retweets from local Starbucks United twitter accounts in support of Palestine and condemning Israel were circulating. On October 13th, Starbucks sent a letter to Workers United, demanding the union to stop using the Starbucks name and similar logo. Workers United retaliated saying that Starbucks was using the tragedy in the Middle East to support their anti-union campaign.

# Data {#sec-data}

A sense of the dataset should be communicated to the reader. The broader context of the dataset should be discussed. All variables should be thoroughly examined and explained. Explain if there were similar datasets that could have been used and why they were not. If variables were constructed then this should be mentioned, and high-level cleaning aspects of note should be mentioned, but this section should focus on the destination, not the journey. It is important to understand what the variables look like by including graphs, and possibly tables, of all observations, along with discussion of those graphs and the other features of these data. Summary statistics should also be included, and well as any relationships between the variables. If this becomes too detailed, then appendices could be used. Basically, for every variable in your dataset that is of interest to your paper there needs to be graphs and explanation and maybe tables.

## Source Data and Methodology

The source data was obtained from Yahoo Finance Canada [@yahoo] and the R package [@R-tidyquant]. The Yahoo Finance Canada historical data was filtered for the desired date range and daily data before being downloaded. The original data had “Date”, “Open”, “High”, “Low”, “Close”, “Adj Close”, and “Volume”. The only variables of interest were “Date” and “Close” which represented the date of the stock’s price and the price when the market closed respectively.

## Limitations

## Data Cleaning
The data was cleaned using the statistical programming language R [@citeR] as well as library packages `tidyverse` [@R-tidyverse], `tinytex` [@R-tinytex],`ggplot2` [@R-ggplot2], `dplyr` [@R-dplyr], `readr` [@R-readr], `tibble` [@R-tibble], `stringr` [@R-stringr], `kableExtra` [@R-kableExtra], `janitor` [@R-janitor],  `xts` [@R-xts], and `knitr` [@knitr2015], and `arrow` [@R-arrow]. The data was cleaned by selecting the two variables “Date” and “Close”  which are shown in @tbl-rawdataNKE below @tbl-rangeData, shows the number of days observed as part of the boycott in addition to one week before it started. It also indicates the start and end date of each dataset. The largest range was Nike at nearly 7 years compared to the less than 4 months Starbucks has been observed for. Each company also represents a different decade, with Nike in the 1990s, Parks in the 2010s, and most recently Starbucks in the 2020s. 





```{r}
#| label: tbl-rawdataNKE
#| tbl-cap: "Description of Raw Nike Stock Variables"
#| echo: false

rawdataNKE <- data.frame( 
  "Variables" = c( "Date", "Close"),
  "New Name" = c( "Date", "Stock Price"), 
  "Description of Variables" = c("The day of the stock price", "The value of the stock at market close"), 
  "Example Value" = c("1991-07-12", "1.355469")
)

knitr::kable(rawdataNKE, 
      col.names = c("Variables", 
                    "New Name",
                    "Description of Variables" ,
                    "Example Value"),
      booktabs = TRUE,
      linesep = " "
      )
```

```{r}
#| label: tbl-rangeData
#| tbl-cap: "The range of time each company's stock was observed for."
#| echo: false


N_NKE = nrow(cleaned_data_NKE)
N_PRKS = nrow(cleaned_data_PRKS)
N_SBUX = nrow(cleaned_data_SBUX)

Day1_NKE = cleaned_data_NKE[1,1]
Day2_NKE = cleaned_data_NKE[N_NKE,1]

Day1_PRKS = cleaned_data_PRKS[1,1]
Day2_PRKS = cleaned_data_PRKS[N_PRKS,1]

Day1_SBUX = cleaned_data_SBUX[1,1]
Day2_SBUX = cleaned_data_SBUX[N_SBUX,1]


dataALL <- data.frame(
  "Companies" = c("Nike (NKE)", "United Parks and Recreation (PRKS)", "Starbucks (SBUX)"),
  "Days Observed" = c(N_NKE, N_PRKS, N_SBUX),
  "Date Start" = c(Day1_NKE, Day1_PRKS, Day1_SBUX),
  "Date End" = c(Day2_NKE, Day2_PRKS, Day2_SBUX)
)

knitr::kable(dataALL, 
      col.names = c("Companies", 
                    "Days Observed",
                    "Date Start" ,
                    "Date End"),
      booktabs = TRUE,
      linesep = " "
      )
```

## Change in Stock Price
The change in stock price over the length of the boycott is observed in @fig-linegraphNKE, @fig-linegraphPRKS, and @fig-linegraphSBUX. These three graphs show a week before the boycott started to the end of the boycott. @fig-linegraphNKE shows how Nike’s stock price was relatively unstable in the first three years of the boycott, with stock prices rising later in 1991 before dropping slightly and then peaking early in 1993 before descending to a low a year later. A progressive cimb from that point to February of 1997 sees a growth of around 750%. @fig-linegraphPRKS shows that United Parks and Recreation had a steep drop in stock price when ‘Blackfish’ released and had a year of poor stability after before dropping drastically. Afterwards the stock price continues to remain unstable.  @fig-linegraphSBUX graphs the stock price of Starbucks and depicts a steep climb in price in November of 2023 with a peak in early December before a steady climb down over the next 5 months. 

```{r}
#| label: fig-linegraphNKE
#| fig-cap: Changes in Nike Stock Price from One Week Before the Commencement of Labour Boycotts to One Week After Their Pledge to End Poor Working Conditions and Child Labour
#| echo: false

cleaned_data_NKE |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "July 12th, 1991 to May 19th, 1998", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "12 months",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(0,10) + 
  theme_minimal()

```

```{r}
#| label: fig-linegraphPRKS
#| fig-cap: Changes in United Parks and Recreation Stock Price from One Week Before Blackfish was Released to One Week After Their Pledge to End Orca Breeding
#| echo: false

cleaned_data_PRKS |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "July 12th, 2013 to May 17th, 2016", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "12 months",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(15,40) + 
  theme_minimal()

```


```{r}
#| label: fig-linegraphSBUX
#| fig-cap: Changes in Starbucks Stock Price from One Week Before Legal Action Against Workers United to March 28th
#| echo: false

cleaned_data_SBUX |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "October 11th, 2023 to March 28th, 2024 ", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "1 month",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(80,110) + 
  theme_minimal()
```


# Results {#sec-results}

## Difference in Stock Price Between Commencement and End of Boycott
@tbl-starttoendNKE, @tbl-starttoendPRKS, and @tbl-starttoendSBUX summarizes the change in stock price between a week before the boycott commenced and its commencement date as well as the difference between the start and end dates of the boycott. The first date is the stock price a week before the stock price and the second date is the day the boycott began.

@tbl-starttoendNKE shows Nike had a small decrease in stock price from a week before the beginning of the boycott. There is also an increase in stock price from the day the boycott started to the day it ended. Relative to the stock price, the change is very large as it increased by 335.82%. @tbl-starttoendPRKS shows that there was a significant decrease of 55.3% in United Parks and Recreation's stock price from the beginning to end of the boycott. Starbucks was relatively unchanged from the start to the end of the boycott with a decrease of only 2.5% as shown in @tbl-starttoendSBUX. 

```{r}
#| label: tbl-starttoendNKE
#| tbl-cap: Changes in Starbucks Stock Price from One Week Before Legal Action Against Workers United, the the Day it was Announced, to March 28th
#| echo: false


starttoend_NKE <- cleaned_data_NKE |>
  filter(Date == '1991-07-12' | Date == '1991-07-18' | Date == '1998-05-19')

starttoend_NKE <-
  starttoend_NKE |> 
  mutate(Difference = Close - lag(Close))

kable(starttoend_NKE, 
      col.names = c("Date", 
                    "Closing Stock Price",
                    "Difference Between Stock Prices"),
      booktabs = TRUE,
      linesep = ""
)
```

```{r}
#| label: tbl-starttoendPRKS
#| tbl-cap: Changes in United Parks & Recreation Stock Price from One Week Before Blackfish was Released, to the Date it was Released, to their Announcement that they will end their Orca Breeding Program
#| echo: false


starttoendPRKS <- cleaned_data_PRKS |>
  filter(Date == '2013-07-12' | Date == '2013-07-19' | Date == '2016-03-16')

starttoendPRKS <-
  starttoendPRKS |> 
  mutate(Difference = Close - lag(Close)) 

kable(starttoendPRKS, 
      col.names = c("Date", 
                    "Closing Stock Price",
                    "Difference Between Stock Prices"),
      booktabs = TRUE,
      linesep = ""
)
```

```{r}
#| label: tbl-starttoendSBUX
#| tbl-cap: Changes in Starbucks Stock Price from One Week Before Legal Action Against Workers United, the the Day it was Announced, to March 28th
#| echo: false


starttoendSBUX <- cleaned_data_SBUX |>
  filter(Date == '2023-10-11' | Date == '2023-10-18' | Date == '2024-03-28')

starttoendSBUX <-
  starttoendSBUX |> 
  mutate(Difference = Close - lag(Close)) 

kable(starttoendSBUX, 
      col.names = c("Date", 
                    "Closing Stock Price",
                    "Difference Between Stock Prices"),
      booktabs = TRUE,
      linesep = ""
)
```

## Detailed Observation of Change in Stock Price
In @tbl-twoweeksNKEprotest, the daily difference in Nike's stock price folowing the international Global Exchange protest. The largest change is a week after the protest on the 27th with a drop in 5.79% or 35 cents. 
@tbl-twoweeksSBUX shows the daily difference in stock price for two weeks from the announcement of the lawsuit. The stock prices were relatively unaffected immediately following the announcement and November 2nd is responsible for the large spike in price shown in @fig-linegraphSBUX

```{r}
#| label: fig-linegraphNKEfocused
#| fig-cap: Changes in Nike Stock Price from CAIR Boycott to Two Weeks After the Global Exchange Protest in 1997
#| echo: false

cleaned_data_NKE |>
  filter(between(Date, as.Date('1997-03-02'), as.Date('1997-11-01'))) |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "July 12th, 1991 to May 19th, 1998", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "1 month",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(5,9) + 
  theme_minimal()

```

```{r}
#| label: tbl-twoweeksNKE
#| tbl-cap: Changes in Nike Stock Price Two Weeks After Global Exchange Protest on October 18th, 1997
#| echo: false

twoweeksNKE <- 
  cleaned_data_NKE |>
  filter(between(Date, as.Date('1997-10-18'), as.Date('1997-11-01')))
  

twoweeksNKE <-
  twoweeksNKE |> 
  mutate(Difference = Close - lag(Close),
  Percentage_Change = (Close - lag(Close)) / lag(Close) * 100)

kable(twoweeksNKE, 
      col.names = c("Date", 
                    "Closing Stock Price",
                    "Difference",
                    "Percentage Change"),
      booktabs = TRUE,
      linesep = ""
)
```


```{r}
#| label: tbl-twoweeksSBUX
#| tbl-cap: Changes in Starbucks Stock Price Two Weeks From the Lawsuit
#| echo: false

twoweeksSBUX <- 
  cleaned_data_SBUX |>
  filter(between(Date, as.Date('2023-10-18'), as.Date('2023-11-06')))

twoweeksSBUX <-
  twoweeksSBUX |> 
  mutate(Difference = Close - lag(Close),
         Percentage_Change = (Close - lag(Close)) / lag(Close) * 100)

kable(twoweeksSBUX, 
      col.names = c("Date", 
                    "Closing Stock Price",
                    "Difference",
                    "Percentage Change"),
      booktabs = TRUE,
      linesep = ""
)
```


## Comparing with the Market
@fig-djiNKE compares the stock price of Nike and the Dow Jones Index (DJI) [@dowjones]. Dow Jones Index includes Nike among around 30 other companies, and it is clear that they have very similar trajectories throughout the boycott. The only apparent difference is the bump in Nike’s price between 1996 and 1998. Otherwise, both stocks have a steady upward growth.

@fig-spxPRKS compares the stock price of United Parks & Recreation with the S&P 500 Index (SPX). The comparison cannot indicate whether United Parks was doing better than the market based on price, but by comparing the change in price it is clear that overall the market and United Parks were on a fall entering into 2015. However, the steep price drop at United Parks cannot be explained by the market. The gradual decline at the beginning of the boycott also does not resemble the upward trajectory in the market in 2014. 

@fig-ndxSBUX compares Starbucks’ stock price with the Nasdaq 100 Index (NDX). The absence of Starbucks’ price spike in the Nasdaq stock shows that it cannot be attributed to market changes. The following decrease in the Starbucks stock price after mid-November is also inconsistent with Nasdaq’s upwards trend. 


```{r}
#| label: fig-spxPRKS
#| fig-cap: Changes in United Parks and Recreation Stock Price in Comparison with the S&P 500 index
#| echo: false

PRKS_SPX <-
  cleaned_data_SPX |>
  filter(between(date, as.Date('2013-07-19'), as.Date('2016-03-16')))

PRKS_Focus <-
  cleaned_data_PRKS |>
  filter(between(Date, as.Date('2013-07-19'), as.Date('2016-03-16')))

SPX_Stocks <- PRKS_SPX$close
PRKS_Stocks <- PRKS_Focus$Close

PRKS_Date <- PRKS_Focus$Date

# Plot PRKS
plot_prks <- ggplot(PRKS_Focus, aes(x = PRKS_Date, y = PRKS_Stocks)) +
  geom_line(color = "#0000c2") +
  labs(title = "United Parks and Rec Stock Price",
       x = "Date", y = "Stock Price") +
  theme_minimal()

# Plot SPX
plot_spx <- ggplot(PRKS_SPX, aes(x = PRKS_Date, y = SPX_Stocks)) +
  geom_line(color = "#FF5733") +
  labs(title = "S&P 500 Index",
       x = "Date", y = "Index Value") +
  theme_minimal()

# Arrange plots
grid.arrange(plot_prks, plot_spx, ncol = 1)
```


```{r}
#| label: fig-djiNKE
#| fig-cap: Changes in the Nike Stock Price in Comparison with the Dow Jones Index
#| echo: false

NIKE_DJI <-
  cleaned_data_DJI |>
  filter(between(date, as.Date('1991-07-18'), as.Date('1998-05-19')))

NIKE_Focus <-
  cleaned_data_NKE |>
  filter(between(Date, as.Date('1991-07-18'), as.Date('1998-05-19')))

DJI_Stocks <- NIKE_DJI$close
NIKE_Stocks <- NIKE_Focus$Close

NIKE_Date <- NIKE_Focus$Date

# Plot Nike
plot_nike <- ggplot(NIKE_Focus, aes(x = NIKE_Date, y = NIKE_Stocks)) +
  geom_line(color = "#0000c2") +
  labs(title = "Nike Stock Price",
       x = "Date", y = "Stock Price") +
  theme_minimal()

# Plot DJI
plot_dji <- ggplot(NIKE_DJI, aes(x = NIKE_Date, y = DJI_Stocks)) +
  geom_line(color = "#FF5733") +
  labs(title = "Dow Jones Index",
       x = "Date", y = "Index Value") +
  theme_minimal()

# Arrange plots
grid.arrange(plot_nike, plot_dji, ncol = 1)

```

```{r}
#| label: fig-ndxSBUX
#| fig-cap: Changes in the Starbucks Stock Price in Comparison with the Nasdaq 100 Index from October 11th, 2023 to March 28th, 2024
#| echo: false

SBUX_NDX <-
  cleaned_data_NDX |>
  filter(between(date, as.Date('2023-10-11'), as.Date('2024-03-28')))

SBUX_Focus <-
  cleaned_data_SBUX |>
  filter(between(Date, as.Date('2023-10-11'), as.Date('2024-03-18')))

NDX_Stocks <- SBUX_NDX$close
SBUX_Stocks <- SBUX_Focus$Close

SBUX_Date <- SBUX_Focus$Date

# Plot SBUX
plot_SBUX <- ggplot(SBUX_Focus, aes(x = SBUX_Date, y = SBUX_Stocks)) +
  geom_line(color = "#0000c2") +
  labs(title = "SBUX Stock Price",
       x = "Date", y = "Stock Price") +
  theme_minimal()

# Plot NDX
plot_NDX <- ggplot(SBUX_NDX, aes(x = SBUX_Date, y = NDX_Stocks)) +
  geom_line(color = "#FF5733") +
  labs(title = "Nasdaq 100 Index",
       x = "Date", y = "Index Value") +
  theme_minimal()

# Arrange plots
grid.arrange(plot_SBUX, plot_NDX, ncol = 1)
```

# Discussion {#sec-discussion}

## Nike: Seemingly unaffected
Nike is the only company with an increase in stock price over the boycott period and it is quite significant at 334%. Despite this overall growth, Nike does not remain unaffected by the boycotts. In 1997 the company suffered a significant decrease in comparison to the overall growth in the market as seen in @graph-djiNKE. 

## SeaWorld: The Tainted Consumer Experience

## Starbucks: Importance of Social Media Among Youth

On November 2nd, Starbucks released their Q4 and Full Year Fiscal Results. Some highlights were record net revenues, store sales up by 8% globally, and record number of stores [@investstarbucks]. In @tbl-twoweeksSBUX, the largest difference in stock price happened on November 2nd at a jump of \$8.67. From observing the comparison with the Nasdaq 100 Index in @fig-ndxSBUX, it is clear that this spike is unrelated to general market trends. This indicates that investors perceived the beginning of the boycott and the lawsuit as harmless to the financial stability and growth of Starbucks. Instead, the positive earnings call had a significant impact on the perceived worth of Starbucks. After the spike, Starbucks began to see possible effects of the boycott. Unlike the upward trending market seen in @fig-ndxSBUX, Starbucks suffered from a gradual decline. At the same time, social media exposure on the Starbucks boycott continued to gain traction. Throughout November 3,000 TikTok videos were posted using the hashtag “#boycottstarbucks” with over 51 million views in total [@nbcstarbucks]. From September to December at the University of Texas, 3 Starbucks-affiliated cafes had decreased sales by around $20,000 each and by February every store on campus with Starbucks licensing had decreased sales [@utstarbucks]. A Starbucks franchise operator in the Middle East and North Africa region announced that they will be laying off 2000 employees as a result of struggling business during the boycott [@MENAlayoffs]. The continuous use of social media to spread awareness on the boycott has undoubtedly impacted the Starbucks business and also brought their stock price down from its peak in November to below its price before the boycott as seen in @tbl-starttoendSBUX. These impacts indicate the role of social media in modern boycotts and its significance in action particularly among youth. The influence of the earnings call indicates that regardless of media attention, if the company’s financial prospects and net income are still high, the stock prices will not be affected.

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References
